---
title: "ACHD Report test"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(lubridate)
library(ggplot2)
library(rgdal)
library(sp)
library(sf)
library(leaflet)
library(leaflet.extras)
library(tableHTML)
```

This report was generated on `r Sys.Date()` from the Adult Congenital Heart Disease Clinic Planning Tool

## Patient Data

``` {r}
beth <- list('1' ="Simple",
             '2' ="Moderate",
             '3' ="Severe",
             '4' = "unknown")

beth.selected <- beth %in% c('1','2','3')

beth.selected

beth[c('1','2','3')]
```

The Patient Dataset had the following filters applied:  
  
* __Selected Genders:__ Male, Female, Neither  
* __CHD Disease Severity:__ `r beth[c('1','2','3')]`
* __Age Range:__ 18 yrs to 100 yrs  
* __Time Period:__ 01/01/2000 to 31/12/2020  
* __Years Since Last Clinic Visit:__ 0 yrs to 21 yrs  
* __Mortality Status:__ Alive  
  

After Filtering, these are some basic demographics of the patient data:  

* __Number of Patients:__ 1000  
* __Number with Simple CHD:__  1000  
* __Number with Moderate CHD:__ 1000
* __Number with Complex CHD:__ 1000
* __Number of Patients Lost to Follow up:__ 1000  
* __Number of Patients with a 1 hour drive to a clinic:__ 1000  




