---
title: "test"
author: "Calum Nicholson"
date: "17/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(shiny)
library(rgdal)
library(sp)
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(ggplot2)
library(sjmisc)
```

``` {r message=FALSE, warning=FALSE, include=FALSE}
# Path to ACHD database data (note: only accessible at RPAH)
pt_data <- 'Z:/CURRENT_STUDIES/2020_achd_map'

# Path to the study's data folder (note: only accessible at RPAH)
app_data <- '../ACHD_Dashboard/data/'

#import ACHD data
achd <- readRDS(file = paste(pt_data, 'output/rpah_analysis_dataset.rds', sep=""))

#import DX coding
dx_codes <- read.csv(file = paste(app_data, '/ACHD_EPCC_coding.csv', sep=""), fileEncoding="UTF-8-BOM") %>% 
            filter(!(Variable == 'adm_AbsentPA' | Variable == 'PA' | Variable == "achd_id")) %>%
            select(!c(ACHD_Database_name, check., Variable))

#import Census data
sa2.TB <- readRDS(file = paste(app_data, 'sa2_demographics.rds', sep="")) %>%
                  mutate(IRSD = ordered(IRSD, c(1,2,3,4,5,6,7,8,9,10)),
                         IRSAD = ordered(IRSAD, c(1,2,3,4,5,6,7,8,9,10)))
```


```{r}
filter.data <- function(data,
                          sb.mortality,
                          sb.sex,
                          sb.bethesda,
                          sb.age,
                          sb.dates,
                          sb.last.clinic) 
       {
        data %>% 
            filter(death %in% sb.mortality) %>%
            filter(sex %in% sb.sex) %>%
            filter(bethesda_code %in% sb.bethesda) %>%
            filter(as.numeric(age, 'years') >= sb.age[1]) %>%
            filter(as.numeric(age, 'years') <= sb.age[2]) %>%
            filter(as.numeric(gap_2000, 'years') >= sb.last.clinic[1]) %>%
            filter(as.numeric(gap_2000, 'years') <= sb.last.clinic[2]) %>%
    
            {if( !(sb.dates[1] == "2000-01-01" & sb.dates[2] == "2020-12-31") )
              
              mutate(., clinics_2000 = map(clinics_2000, ~ .x %>%
                                        mutate(
                                          clinic_in_period = 
                                            sapply(.$clinic_date, 
                                                   function(x) x %within% (as.Date(sb.dates[1])
                                                                          %--% 
                                                                          as.Date(sb.dates[2]))))),
                   in_time_period = map_dbl(map(clinics_2000, ~ .$clinic_in_period), any)
                   ) %>%
              filter(in_time_period == TRUE)
              
              else . }
       }
```

``` {r}
sb.mortality <- c(0)
sb.sex <- c(1,2,4)
sb.bethesda <- c(1,2,3)
sb.age <- c(18,110)
sb.dates <- c("2000-01-01", "2020-12-31")
sb.last.clinic <- c(10, 21)

achd.filtered <- filter.data(achd,
                             sb.mortality,
                             sb.sex,
                             sb.bethesda,
                             sb.age,
                             sb.dates,
                             sb.last.clinic)
                
```

``` {r}
print("Mortality")
unique(achd.filtered$death)

print("Sex")
unique(achd.filtered$sex)

print("Bethesda")
unique(achd.filtered$bethesda_code)

print("Age")
range(as.numeric(achd.filtered$age, "years"))

```

``` {r}
input <- 2


if (input == 2) {
            achd.df <- achd %>% dplyr::rename("area" = sa2)
            TB.df <- sa2.TB %>% dplyr::rename("area" = SA2_NAME)
} else if (input == 3) {
            achd.df <- achd %>% dplyr::rename("area" = sa3)
} else {
            achd.df <- achd %>% dplyr::rename("area"= sa4)}

area.achd <- achd.df %>%
              group_by(area) %>%
              dplyr::summarise(ACHD_count = n(), 
                        beth_1 = sum(bethesda_code == 1),
                        beth_2 = sum(bethesda_code == 2), 
                        beth_3 = sum(bethesda_code == 3), 
                        beth_4 = sum(bethesda_code == 4))

area.dx <- achd.df %>%
            mutate_at(as.character(dx_codes[['EPCC_Code']]), as.character) %>% 
            mutate_at(as.character(dx_codes[['EPCC_Code']]), as.numeric) %>%
            group_by(area) %>% 
            dplyr::summarise_at(as.character(dx_codes[['EPCC_Code']]), sum, na.rm = TRUE)


area.data <- left_join(TB.df, area.achd, by = 'area') %>% 
             left_join(area.dx, by = 'area') %>%
                mutate_at(as.character(dx_codes[['EPCC_Code']]), function(x) replace(x, is.na(x), 0)) %>%
                mutate_at(c('ACHD_count', 'beth_1', 'beth_2', 'beth_3', 'beth_4'), function(x) replace(x, is.na(x), 0))
                          
```

``` {r}
area.dx <- area.data %>% filter(SA2_5DIGIT == 11007) %>% select(dx_codes$EPCC_Code) %>% 
            t() %>% as.data.frame() %>% 
            rename("dx_count" = V1) %>% 
            mutate(dx_label = dx_codes$Adjust_name[match(row.names(.), dx_codes$EPCC_Code)])

```
``` {r}
ggplot(area.dx, aes(x = dx_label, y=dx_count)) + 
                    geom_bar(stat="identity") +
                    theme(axis.text.x = element_text(angle = 90)) +
                    labs(title = "Frequecy of each diagnosis", 
                         y = "count",
                         x = "") +
                    coord_flip()
```

``` {r}

test <- area.data %>% select(area, SA2_5DIGIT, IRSD, male) %>%
                      mutate(male = replace(male, male == Inf, NA))


test$cuts <- cut(as.numeric(test$IRSD), breaks = 10, label = F)
#coloured_cut <- test$cuts[test$SA2_5DIGIT == 11007]
test$color <- ifelse(test$cuts == test$cuts[test$SA2_5DIGIT == 11007][1], "Selected", "")

test %>% filter(!is.na(IRSD)) %>%
          ggplot(aes(x=IRSD, fill = cuts == test$cuts[test$SA2_5DIGIT == 11007][1])) +
                geom_bar() +
                scale_fill_manual(values = c("grey45", "red"))+
                theme_minimal() + theme(legend.position = "none")
```

``` {r}
library(scales)

rem.test <- area.data %>% select(area, SA2_5DIGIT, major_cities:very_remote) 

  datalong <- rem.test %>% filter(SA2_5DIGIT == 11007) %>%
                pivot_longer(cols = major_cities:very_remote, names_to = "remoteness")

datalong$remoteness <- factor(datalong$remoteness, levels = datalong$remoteness)

ggplot(data = datalong, aes(x = factor(remoteness), y = value)) +
  geom_col() +
  scale_y_continuous(limits = c(0,1), labels = percent) +
  theme_minimal()

```

``` {r}
achd %>% mutate(gap = as.duration(as.Date('01/10/2020', format='%d/%m/%Y') - last_clinic_2000) / dyears(1),
                ltf_3 = ifelse(gap >= 3, 1, 0),
                ltf_4 = ifelse(gap >= 4, 1, 0),
                ltf_5 = ifelse(gap >= 5, 1, 0),
                ) %>%
         select(last_clinic_2000, gap, ltf_3, ltf_4, ltf_5)

```


``` {r}
achd.1 <- achd %>% mutate(gap = as.duration(as.Date('01/10/2020', format='%d/%m/%Y') - last_clinic_2000) / dyears(1),
                          ltf_3 = ifelse(gap >= 3, 1, 0),
                          ltf_4 = ifelse(gap >= 4, 1, 0),
                          ltf_5 = ifelse(gap >= 5, 1, 0),
                          ) %>%
                   select(sa4, bethesda_code, ltf_3, ltf_4, ltf_5)


achd.1 %>% group_by(sa4) %>% dplyr::summarise(n = n(), 
                                              beth_1 = sum(bethesda_code == 1), 
                                              beth_2 = sum(bethesda_code == 2),
                                              beth_3 = sum(bethesda_code == 3),
                                              beth_4 = sum(bethesda_code == 4),
                                              ltf_3 = sum(ltf_3),
                                              ltf_4 = sum(ltf_4),
                                              ltf_5 = sum(ltf_5),
                                              )
````





```{r}
polys <- readOGR('../ACHD_Dashboard/data/shape_files/sa2_polys.shp')
```



``` {r}
polys$CODE16[polys$NAME16 == 'Riverina']

polys$NAME16[polys$CODE16 == 113]
```

``` {r}
# add the achd population for each area 
polys <- merge(polys, area.achd, by.x = 'NAME16', by.y = 'area')
#convert NAs to 0
polys@data <- polys@data %>% replace_na(list('ACHD_count' = 0))

```


``` {r}
id = as.numeric('113')


polys$CODE16[polys$NAME16 == 'Riverina']

polys$NAME16[polys$CODE16 == id]
```


# Selecting New Clinics
                  tabBox(title = NULL,
                         tabPanel("Select Hospitals by location",
                                  uiOutput('phn.selector'), # Primary Health Network
                                  uiOutput('lhn.selector'), # Local Health Network
                                  uiOutput('hospital.selector'), # Hospital
                                  div(style="display:inline-block",uiOutput('clinic.button.add')), # Add hospital to list
                                  div(style="display:inline-block",uiOutput('clinic.button.reset')), # Reset list
                                  ),
                          tabPanel("Filter Hospitals by Type",
                                   uiOutput('description.filter'),
                                   uiOutput('sector.filter'),
                                   uiOutput('beds.filter'),
                                   div(style="display:inline-block",uiOutput('clinic.button.add')), # Add hospital to list
                                   div(style="display:inline-block",uiOutput('clinic.button.reset')), # Reset list
                          ),
                        width = 4, height = 340
                  ),
                  
                  
                  box(title = "Select Hospitals by location",
                                  uiOutput('phn.selector'), # Primary Health Network
                                  uiOutput('lhn.selector'), # Local Health Network
                                  uiOutput('hospital.selector'), # Hospital
                                  div(style="display:inline-block",uiOutput('clinic.button.add')), # Add hospital to list
                                  div(style="display:inline-block",uiOutput('clinic.button.reset')), # Reset list
                        width = 4, height = 340
                  ),


``` {r}
ids <- c(1, 2, 3, 4, 5, 6, 7)

new <- c(5, 6, 7, 8, 9, 10)
```


``` {r}
any(new %in% ids)

new[!(new %in% ids)]


```






