---
title: "Table Builder EDA"
author: "Calum Nicholson"
date: "08/06/2020"
output: html_document
---

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# set 'data' folder as root directory
knitr::opts_knit$set(root.dir = '../..')

library(tidyverse)
library(lubridate)
```

## Function For importing Table Builder data 

Table Builder data was downloaded on 5 Jun 2020. The data is located in Dissertation/data/Table_builder_SA2_NSW.

Each table row in this folder is an SA2 area for NSW, with the columns describing a certain demographic. This function will import the data and remove the unneccesary rows.

``` {r}
getwd()

```

``` {r}
# Function to import and clean SA2 Tables
read.SA2.table <-function (filename, path = "./Table_builder_SA2/") {
  
#    ```
#    Takes an table downloaded from Table Builder at the SA2 levels, containing all NSW SA2 #    areas, and converts it to a dataframe
#    
#    INPUT
#    filename:       string, name of the file with extension
#    path:           string, path to the file
#    
#    OUTPUT
#    sa2.df:         dataframe, Table builder data with excess removed
#    ```
  
    #load table
    table <- readLines(paste(path,filename, sep=''))
    
    #tidy the data
    table <- table[-(1:9)] #remove header
    table <- table[-2] #remove SA2 label
    
    #load data into dataframe
    sa2.df <- read.table(text = table, sep =",", header = TRUE, stringsAsFactors = FALSE, fill = TRUE)
    
    sa2.df <- head(sa2.df, -4) %>% #remove footer
              select(-X)
    
    #convert all columns except header into numeric
    sa2.df[-1] <- lapply(sa2.df[-1], as.numeric)
    
    sa2.df
}

head(read.SA2.table('age_10P.csv'))
     
```

## Creating Area level variables

Going though each of the tables, adding variables into a final dataset, sa2.data.

__total population, sex and country of birth__
``` {r}

# load some tables
sex <- read.SA2.table('sex.csv')
COB <- read.SA2.table('COB.csv')

# SA2 area names
sa2_area <- sex$SEXP.Sex

# Total population in each area
total_pop <- sex$Total

# Percentages in each area
male <- ( sex$Male / total_pop ) * 100
female <- ( sex$Female / total_pop ) * 100
overseas <- ( COB$COB_overseas / total_pop ) * 100

# combine variables
sa2.data <- data.frame(sa2_area, total_pop, male, female, overseas)
head(sa2.data)

```
__childcare__
``` {r}
# load data
childcare_cat <- read.SA2.table('child_care.csv')

# percentage of population engaged in childcare
childcare <-  childcare_cat %>%
                mutate(sum = rowSums(.[3:5])) %>%
                mutate(perc = sum / Total * 100) %>%
                select(CHCAREP.Unpaid.Child.Care, perc)

# update column names
names(childcare) <- c('sa2_area','childcare')

# add to data set
sa2.data2 <- left_join(sa2.data, childcare, by = 'sa2_area')

``` 

__employment__
``` {r}
# load data
employ_cat <- read.SA2.table('employment_status.csv')

# percentage of unemployed in each area
employ <- employ_cat %>%
              mutate(sum = rowSums(.[5:6])) %>%
              mutate(perc = sum / Total * 100) %>%
              select(LFSP.Labour.Force.Status, perc)

# update column names
names(employ) <- c('sa2_area','employ')

# add to data set
sa2.data3 <- left_join(sa2.data2, employ, by = 'sa2_area')

``` 

__overcrowding__
``` {r}
# load data
overcrowding_cat <- read.SA2.table('overcrowding.csv')

# Overcrowding measured by number of spare rooms, or number of rooms needed in household
# Percentage who have spare rooms (positive), and percentage who need more rooms (negative)
overcrowding <- overcrowding_cat %>%
              mutate(positive = rowSums(.[7:10])) %>%
              mutate(negative = rowSums(.[2:5])) %>%
              mutate(perc_pos = positive / Total * 100) %>%
              mutate(perc_neg = negative / Total * 100) %>%
              select(HOSD.Housing.Suitability, perc_pos, perc_neg)

# update column names
names(overcrowding) <- c('sa2_area','oc_positive', 'oc_negative')

# add to data set
sa2.data4 <- left_join(sa2.data3, overcrowding, by = 'sa2_area')

```

__single parents__
``` {r}
# load data
single_parent_cat <- read.SA2.table('single_parent.csv')

#Percentage of single parent households; male, female and total
single_parent <- single_parent_cat %>%
              mutate(sp_male = Male.lone.parent / Total * 100) %>%
              mutate(sp_female = Female.lone.parent / Total * 100) %>%
              mutate(sp_total = sp_male + sp_female) %>%  
              select(SLPP.Sex.of.Lone.Parent, sp_male, sp_female, sp_total)

# update column names
names(single_parent) <- c('sa2_area','sp_male','sp_female','sp_total')

# add to data set
sa2.data5 <- left_join(sa2.data4, single_parent, by = 'sa2_area')
```

__students__
``` {r}
# load data
student_cat <- read.SA2.table('student_status.csv')

# percentage of students in each area
student <- student_cat %>%
              mutate(student_sum = rowSums(.[3:4])) %>%
              mutate(student_perc = student_sum / Total * 100) %>% 
              select(STUP.Full.Time.Part.Time.Student.Status, student_perc)

# update column names
names(student) <- c('sa2_area','student')

# add to data set
sa2.data6 <- left_join(sa2.data5, student, by = 'sa2_area')


```

__age__
``` {r}
# load data
#add age groups for each area
age_cat <- read.SA2.table('age_10P.csv')

# age categories, in 10 year age groups.
age_cat <- select(age_cat, AGE10P...Age.in.Ten.Year.Groups:X100.years.and.over)

# update column names
names(age_cat) <- c('sa2_area','age0','age10',
                    'age20','age30','age40',
                    'age50','age60','age70',
                    'age80','age90','age100')

# add to data set
sa2.data7 <- left_join(sa2.data6, age_cat, by = 'sa2_area')

```

__Disadvantage__
``` {r}
irsd <- read.SA2.table('seifa_irsd.csv')

irsd_cat <- irsd %>% select(starts_with('Decile'),IRSD.Deciles.at.SA2.Level..Area.) %>%
                gather(type, value, -IRSD.Deciles.at.SA2.Level..Area.) %>%
                filter(value > 0) %>%
                select(-value) %>%
                rename('sa2_area' = IRSD.Deciles.at.SA2.Level..Area.,
                       'IRSD' = type) %>%
                filter(sa2_area != 'Total')


irsd_cat$IRSD <- gsub("Decile.", "", irsd_cat$IRSD)
irsd_cat$IRSD <- as.factor(irsd_cat$IRSD)


sa2.data8 <- left_join(sa2.data7, irsd_cat, by = 'sa2_area')

```


__Advantage and Disadvantage__
``` {r}
irsad <- read.SA2.table('seifa_irsad.csv')

irsad_cat <- irsad %>% select(starts_with('Decile'),IRSAD.Deciles.at.SA2.Level..Area.) %>%
                gather(type, value, -IRSAD.Deciles.at.SA2.Level..Area.) %>%
                filter(value > 0) %>%
                select(-value) %>%
                rename('sa2_area' = IRSAD.Deciles.at.SA2.Level..Area.,
                       'IRSAD' = type) %>%
                filter(sa2_area != 'Total')

irsad_cat$IRSAD <- gsub("Decile.", "", irsad_cat$IRSAD)
irsad_cat$IRSAD <- as.factor(irsad_cat$IRSAD)

sa2.data9 <- left_join(sa2.data8, irsad_cat, by = 'sa2_area')

```

__Remoteness__
``` {r}
remote <- read.SA2.table('remoteness.csv') %>%
                mutate(major_cities = Major.Cities.of.Australia..NSW. + Major.Cities.of.Australia..ACT.,
                       inner_regional = Inner.Regional.Australia..NSW. + Inner.Regional.Australia..ACT.) %>%
                rename("outer_regional" = Outer.Regional.Australia..NSW.,
                       "remote" = Remote.Australia..NSW.,
                       "very_remote" = Very.Remote.Australia..NSW.,
                       "sa2_area" = RA)%>%
                select(sa2_area, major_cities, inner_regional, outer_regional, remote, very_remote, Total)

remote_perc <- remote %>%
                mutate(major_cities = major_cities / Total,
                       inner_regional = inner_regional / Total,
                       outer_regional = outer_regional / Total,
                       remote = remote / Total,
                       very_remote = very_remote / Total) %>%
                select(sa2_area, major_cities, inner_regional, outer_regional, remote, very_remote)

sa2.data10 <- left_join(sa2.data9, remote_perc, by = 'sa2_area')


```

``` {r}
library(ggplot2)
library(scales)

sa2 = "Total"

datalong <- remote_perc %>% filter(sa2_area == sa2) %>%
                pivot_longer(cols = major_cities:very_remote, names_to = "remoteness")

datalong$remoteness <- factor(datalong$remoteness, levels = datalong$remoteness)

ggplot(data = datalong, aes(x = factor(remoteness), y = value)) +
  geom_col() +
  scale_y_continuous(limits = c(0,1), labels = percent)



```

__Add SA Codes to dataset__
``` {r}
# Load the list of ASGS SA2 area names and codes
SA2_aus <- read.csv('./ASGS/SA2_2016_AUST.csv')

# Filter for NSW and select onyl SA2 associated columns
SA2_nsw <- SA2_aus %>%
            filter(STATE_NAME_2016 == "New South Wales" | STATE_NAME_2016 == "Australian Capital Territory") %>%
            select(SA2_MAINCODE_2016, SA2_5DIGITCODE_2016, SA2_NAME_2016,
                   GCCSA_NAME_2016) %>%
            rename("GCC_NAME" = GCCSA_NAME_2016,
                   "SA2_5DIGIT" = SA2_5DIGITCODE_2016,
                   "sa2_area" = SA2_NAME_2016)

# Add the SA2 area codes to the new dataset
sa2.data11 <- left_join(sa2.data10, SA2_nsw, by = 'sa2_area')

```

__Hospital Travel Time Data__

This data contains the distance between each SA2 area in australia and each hospital in australia. This obviously contains much more data than we need, so we can filter it to look at only SA2 areas in NSW and NSW hospitals with an ACHD service.

``` {r}
# import the hospital travel times data
htt <- read.csv('./hospital_TT/duration_sa2_hospitals.csv') %>%
#         ------------Clean up the data-------------------------  
          # select columns, sa2 code and ACHD locations          
          select(SA2_5DIG16, time_to_646, time_to_755, time_to_152, 
                 time_to_683, time_to_737, time_to_979) %>%
          # rename locations to something understandable
          dplyr::rename("time_to_orange" = time_to_646,
                 "time_to_nowra" = time_to_755,
                 "time_to_canberra" = time_to_152,
                 "time_to_port" = time_to_683,
                 "time_to_rpa" = time_to_737,
                 "time_to_westmead" = time_to_979,
                 "SA2_5DIGIT" = SA2_5DIG16) %>%
          # calculate the shortest travel time to an achd clinic
          mutate(shortest_time = pmin(time_to_orange,time_to_nowra,time_to_canberra,
                                     time_to_port,time_to_rpa,time_to_westmead))

htt.1 <- htt %>%
          mutate(closest_clinic = colnames(htt)[apply(htt,1,which.min)])
```

``` {r}
#merge the travel time data with the area level dataset
sa2.data12 <- left_join(sa2.data11, htt, by = 'SA2_5DIGIT') %>%
             # Convert the numeric data to duration data (in seconds)
             mutate(time_to_orange = duration(time_to_orange, "seconds"),
                    time_to_nowra = duration(time_to_nowra, "seconds"),
                    time_to_canberra = duration(time_to_canberra, "seconds"),
                    time_to_port = duration(time_to_port, "seconds"),
                    time_to_rpa = duration(time_to_rpa, "seconds"),
                    time_to_westmead = duration(time_to_westmead, "seconds"),
                    shortest_time = duration(shortest_time, "seconds"),)
```

``` {r}
# Final table
knitr::kable(head(sa2.data12))

#save as csv and r data file
write.csv(sa2.data12, "./achd-dissertation/analysis_pipeline/output/sa2_table_builder.csv")
saveRDS(sa2.data12, file = "./achd-dissertation/analysis_pipeline/output/sa2_table_builder.rds")

```

``` {r}
test <- readRDS(file = "./achd-dissertation/analysis_pipeline/output/sa2_table_builder.rds")
```








