---
title: "05_ACHD_database"
author: "Calum Nicholson"
date: "03/04/2021"
output: html_document
---

__To do__
Select data for patients with correct address recorded  
Check through values for clinic_comment, what do they all mean how can we organise them? 
Check how removing missing data affects the demographics.



```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Path to ACHD database data (note: only accessible at RPAH)
ACHD_path <- 'Z:CURRENT_STUDIES/2020_achd_map/DATA/'


library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
```

__Create a 'Person-Level' Dataset__

The ACHD database contains a number of tables, joined by the person identified "person_id". These will be joined into a single person level datasat. The key tables are:
    -    ADMIN: names, addresses, contact details, dob, sex.  
    -    MAIN: contains the minimum dataset, with diagnosis information and who their clincian is (i,e. DC, RC or SHC).  
    -    DEMOGRAPHICS: This contains basic demographics information. A lot of this table in incomplete, however there is useful information in country of birth and indigenous status.  
    -    FOLLOW UP STATUS: This is a 'clinic level' dataset, where each record is a clinic visit, with the date of the visit and a comment about which clinic is was (i.e. which private doctor, or public clinic or outreach clinic)     
    -    CLINIC VISIT: Similar to the follow up table but contained more detailed info (i.e bp and heart rate). has only been collecting info for the past two years and only for DCs patients  
    -    DEATH: Recording deceased patients, taken from NDI data linkage.


_ADMIN_
``` {r}
# Load ADMIN table from ACHD database
ADMIN <- read_csv(paste(ACHD_path, 'ADMIN.csv', sep="")) %>%
#         ------------Clean up the data-------------------------  
          # select columns
          select(person_id, adm_sex, adm_date_of_birth,                                   
                 adm_suburb_locality, adm_postcode, adm_state_abbreviation) %>%
          # rename columns
          rename('achd_id' = person_id,                                                   
                 'sex' = adm_sex,
                 'dob' = adm_date_of_birth,
                 'suburb' = adm_suburb_locality,
                 'postcode' = adm_postcode,
                 'state' = adm_state_abbreviation) %>%
          # remove timestamp from visit date
          mutate_at("dob", str_remove, pattern = " 0:00:00") %>%
          # Convert clinic dates to Date format
          mutate_at("dob", as.Date, format='%d/%m/%Y') %>% 
          # Calculate age on 01 Jan 2020
          mutate(age = as.period(dob %--% as.Date("01/01/2020", format='%d/%m/%Y'))) %>%
          # Flag for ages greater than 18 years old
          mutate(age_flag = ifelse(as.numeric(age, "years") >= 18, 1, 0)) %>%
          # Remove whitespace from suburb name
          mutate_at("suburb", str_trim, side = "both") %>% 
          # Change all suburbs to uppercase
          mutate_at("suburb", str_to_upper, locale = "en") %>% 
          # Convert suburb from to character format
          mutate_at("suburb", as.character) %>%     
          # Convert postcode to character format
          mutate_at("postcode", as.character)                                              
```

__Selecting for records with valid addresses__

Check the dataset for missing postcodes, states, overseas or interstate addresses. a few of the records have some missing data but we can safely infer what the missing information is and fix them up:  
      -    achd_id 910 has subrub of CLAREVILLE, but is missing the postcode, CLAREVILLE only matches to one SA2 area in the lookup table so we can safely assume the postcode is the same as the lookup table: 2107  
      -    achd_ids 11, 1385, 2792m, 2034 4023, 4732, 4883, 4892 the state field missing, but the postcode and suburb are in NSW, so we can add the state as NSW  


``` {r}
ADMIN <- ADMIN %>%
                mutate(
                  postcode = replace(postcode, achd_id == 910, 2107),
                  state = replace(state, achd_id == 4732, "NSW"),
                  state = replace(state, achd_id == 4883, "NSW"),
                  state = replace(state, achd_id == 4892, "NSW"),
                  state = replace(state, achd_id == 1385, "NSW"),
                  state = replace(state, achd_id == 2792, "NSW"),
                  state = replace(state, achd_id == 2034, "NSW"),
                  state = replace(state, achd_id == 11, "NSW"),
                  state = replace(state, achd_id == 4023, "NSW"),
                       ) %>%
                mutate(
                  valid_address = ifelse( ( suburb != "" & suburb != "OVERSEAS" & suburb != "-") &
                                            (postcode != "" & !is.na(postcode)) &
                                            ( state == "NSW" | state == "ACT"), 1, 0)
                       )
```

There are some typos in the address fields from the dataset, resulting in 230 records that need to be checked by hand.

__Cleaning Addresses__

The addresses above have some typos, they will be cross references with the sa2 lookup table, first checking the suburb, then the postcode to see if an obvious typo can be identified. If there is no reasonable solution, the record will be marked as an invalid address

``` {r}
ADMIN <- ADMIN %>%
                mutate(
                  postcode = replace(postcode, achd_id == 4688, 2112),
                  suburb = replace(suburb, achd_id == 4675, 'SYDNEY'),
                  postcode = replace(postcode, achd_id == 327, 2203),
                  valid_address = replace(valid_address, achd_id == 3360, 0),
                  postcode = replace(postcode, achd_id == 2672, 2006),
                  postcode = replace(postcode, achd_id == 2065, 2047),
                  postcode = replace(postcode, achd_id == 2120, 2046),
                  postcode = replace(postcode, achd_id == 1361, 2006),
                  postcode = replace(postcode, achd_id == 34, 2045),
                  postcode = replace(postcode, achd_id == 2524, 2198),
                  suburb = replace(suburb, achd_id == 4107, 'DULWICH HILL'),
                  postcode = replace(postcode, achd_id == 444, 2066),
                  suburb = replace(suburb, achd_id == 2382, 'TORONTO'),
                  suburb = replace(suburb, achd_id == 156, 'MOUNT KURING-GAI'),
                  postcode = replace(postcode, achd_id == 256, 2017),
                  suburb = replace(suburb, achd_id == 406, 'NORTH ENTRANCE BEACH'),
                  postcode = replace(postcode, achd_id == 81, 2049),
                  suburb = replace(suburb, achd_id == 275, 'JAMISON CENTRE'),
                  postcode = replace(postcode, achd_id == 1259, 2615),
                  suburb = replace(suburb, achd_id == 4894, 'HELENSBURGH'),
                  suburb = replace(suburb, achd_id == 78, 'ROBERTSON'),
                  postcode = replace(postcode, achd_id == 1583, 2193),
                  suburb = replace(suburb, achd_id == 2695, 'SYLVANIA'),
                  suburb = replace(suburb, achd_id == 2639, 'WAVERLY'),
                  valid_address = replace(valid_address, achd_id == 2302, 0),
                  postcode = replace(postcode, achd_id == 2019, 2478),
                  postcode = replace(postcode, achd_id == 2871, 2040),
                  postcode = replace(postcode, achd_id == 2388, 2006),
                  postcode = replace(postcode, achd_id == 1973, 2006),
                  postcode = replace(postcode, achd_id == 254, 2006),
                  suburb = replace(suburb, achd_id == 1445, 'FIVE DOCK'),
                  suburb = replace(suburb, achd_id == 4995, 'NORTHBRIDGE'),
                  valid_address = replace(valid_address, achd_id == 906, 0),
                  valid_address = replace(valid_address, achd_id == 4337, 0),
                  valid_address = replace(valid_address, achd_id == 539, 0),
                  suburb = replace(suburb, achd_id == 911, 'TRUNDLE'),
                  postcode = replace(postcode, achd_id == 4001, 2825),
                  suburb = replace(suburb, achd_id == 5054, 'BERKELEY VALE'),
                  suburb = replace(suburb, achd_id == 5042, 'DUBBO'),
                  suburb = replace(suburb, achd_id == 1861, 'CROYDON'),
                  postcode = replace(postcode, achd_id == 26, 2137),
                  postcode = replace(postcode, achd_id == 4237, 2585),
                  postcode = replace(postcode, achd_id == 318, 2006),
                  postcode = replace(postcode, achd_id == 1094, 2204),
                  suburb = replace(suburb, achd_id == 4545, 'GREEN VALLEY'),
                  postcode = replace(postcode, achd_id == 3691, 2006),
                  valid_address = replace(valid_address, achd_id == 4585, 0),
                  postcode = replace(postcode, achd_id == 3390, 2261),
                  postcode = replace(postcode, achd_id == 4272, 2212),
                  postcode = replace(postcode, achd_id == 4623, 2166),
                  valid_address = replace(valid_address, achd_id == 2514, 0),
                  postcode = replace(postcode, achd_id == 2419, 2617),
                  postcode = replace(postcode, achd_id == 2551, 2006),
                  valid_address = replace(valid_address, achd_id == 1263, 0),
                  suburb = replace(suburb, achd_id == 1053, 'DALEYS POINT'),
                  valid_address = replace(valid_address, achd_id == 1164, 0),
                  suburb = replace(suburb, achd_id == 1672, 'BELLINGEN'),
                  suburb = replace(suburb, achd_id == 4917, 'ERSKINEVILLE'),
                  valid_address = replace(valid_address, achd_id == 4246, 0),
                  suburb = replace(suburb, achd_id == 876, 'VILLAWOOD'),
                  postcode = replace(postcode, achd_id == 3605, 2913),
                  suburb = replace(suburb, achd_id == 876, 'LONDONDERRY'),
                  postcode = replace(postcode, achd_id == 2870, 2006),
                  postcode = replace(postcode, achd_id == 4923, 2006),
                  suburb = replace(suburb, achd_id == 4615, 'WINDERA'),
                  valid_address = replace(valid_address, achd_id == 1730, 0),
                  valid_address = replace(valid_address, achd_id == 3239, 0),
                  suburb = replace(suburb, achd_id == 860, 'BOWNING'),
                  valid_address = replace(valid_address, achd_id == 593, 0),
                  suburb = replace(suburb, achd_id == 5180, 'MARRICKVILLE'),
                  suburb = replace(suburb, achd_id == 2946, 'WOOLLOOMOOLOO'),
                  valid_address = replace(valid_address, achd_id == 1084, 0),
                  suburb = replace(suburb, achd_id == 3457, 'LONG BEACH'),
                  postcode = replace(postcode, achd_id == 3198, 2045),
                  valid_address = replace(valid_address, achd_id == 4206, 0),
                  valid_address = replace(valid_address, achd_id == 1269, 0),
                  suburb = replace(suburb, achd_id == 5058, 'ST HELENS PARK'),
                  suburb = replace(suburb, achd_id == 3021, 'OXLEY PARK'),
                  suburb = replace(suburb, achd_id == 629, 'LAKE HAVEN'),
                  postcode = replace(postcode, achd_id == 832, 2550),
                  postcode = replace(postcode, achd_id == 1704, 2045),
                  postcode = replace(postcode, achd_id == 1842, 2006),
                  valid_address = replace(valid_address, achd_id == 1924, 0),
                  suburb = replace(suburb, achd_id == 3308, 'KURRI KURRI'),
                  suburb = replace(suburb, achd_id == 1881, 'DARLING POINT'),
                  suburb = replace(suburb, achd_id == 2447, 'BAULKHAM HILLS'),
                  suburb = replace(suburb, achd_id == 2175, 'YAGOONA'),
                  postcode = replace(postcode, achd_id == 2007, 2204),
                  postcode = replace(postcode, achd_id == 1559, 2018),
                  postcode = replace(postcode, achd_id == 2963, 2671),
                  postcode = replace(postcode, achd_id == 1433, 2528),
                  postcode = replace(postcode, achd_id == 907, NA),
                  suburb = replace(suburb, achd_id == 4179, 'ERSKINEVILLE'),
                  suburb = replace(suburb, achd_id == 1942, 'WOLLSTONECRAFT'),
                  postcode = replace(postcode, achd_id == 1594, 2006),
                  suburb = replace(suburb, achd_id == 1446, 'BIDWILL'),
                  suburb = replace(suburb, achd_id == 3343, 'SYLVANIA'),
                  postcode = replace(postcode, achd_id == 5175, 2006),
                  valid_address = replace(valid_address, achd_id == 2605, 0),
                  valid_address = replace(valid_address, achd_id == 1988, 0),
                  suburb = replace(suburb, achd_id == 2688, 'MAYFIELD'),
                  valid_address = replace(valid_address, achd_id == 2281, 0),
                  postcode = replace(postcode, achd_id == 3644, 2006),
                  valid_address = replace(valid_address, achd_id == 1404, 0),
                  postcode = replace(postcode, achd_id == 2919, 2104),
                  suburb = replace(suburb, achd_id == 4252, 'BRADDON'),
                  postcode = replace(postcode, achd_id == 3494, 2006),
                  postcode = replace(postcode, achd_id == 2477, 2161),
                  postcode = replace(postcode, achd_id == 2694, 2115),
                  suburb = replace(suburb, achd_id == 4426, 'RYDE'),
                  postcode = replace(postcode, achd_id == 2550, 2088),
                  suburb = replace(suburb, achd_id == 1470, 'MARRICKVILLE'),
                  suburb = replace(suburb, achd_id == 926, 'CAMDEN'),
                  suburb = replace(suburb, achd_id == 1091, '	WOLLONGONG WEST'),
                  postcode = replace(postcode, achd_id == 4420, 2006),
                  suburb = replace(suburb, achd_id == 1153, 'WOOLLOOMOOLOO'),
                  postcode = replace(postcode, achd_id == 4873, 2536),
                  suburb = replace(suburb, achd_id == 177, 'WISEMANS FERRY'),
                  valid_address = replace(valid_address, achd_id == 2445, 0),
                  suburb = replace(suburb, achd_id == 4509, 'CHIPPING NORTON'),
                  postcode = replace(postcode, achd_id == 1253, 2211),
                  valid_address = replace(valid_address, achd_id == 4486, 0),
                  postcode = replace(postcode, achd_id == 4515, 2528),
                  suburb = replace(suburb, achd_id == 4740, 'BRIGHTON-LE-SANDS'),
                  suburb = replace(suburb, achd_id == 3353, 'WOY WOY'),
                  valid_address = replace(valid_address, achd_id == 1964, 0),
                  suburb = replace(suburb, achd_id == 3695, 'MOUNT PLEASANT'),
                  suburb = replace(suburb, achd_id == 2107, 'WAVERTON'),
                  suburb = replace(suburb, achd_id == 3346, 'MERRYLANDS'),
                  postcode = replace(postcode, achd_id == 1212, 2062),
                  postcode = replace(postcode, achd_id == 3125, 2258),
                  postcode = replace(postcode, achd_id == 1159, 2769),
                  suburb = replace(suburb, achd_id == 1159, 'THE PONDS'),
                  valid_address = replace(valid_address, achd_id == 1802, 0),
                  postcode = replace(postcode, achd_id == 2049, 2155),
                  postcode = replace(postcode, achd_id == 3194, 2023),
                  postcode = replace(postcode, achd_id == 1928, 2259),
                  valid_address = replace(valid_address, achd_id == 1298, 0),
                  suburb = replace(suburb, achd_id == 3070, 'STRATHFIELD'),
                  postcode = replace(postcode, achd_id == 2506, 2006),
                  suburb = replace(suburb, achd_id == 2887, 'STRATHFIELD SOUTH'),
                  postcode = replace(postcode, achd_id == 1128, 2705),
                  suburb = replace(suburb, achd_id == 1244, 'SCHOFIELDS'),
                  postcode = replace(postcode, achd_id == 845, 2010),
                  postcode = replace(postcode, achd_id == 2353, 2233),
                  postcode = replace(postcode, achd_id == 1261, 2258),
                  postcode = replace(postcode, achd_id == 3825, 2234),
                  postcode = replace(postcode, achd_id == 1110, 2160),
                  suburb = replace(suburb, achd_id == 1110, 'MERRYLANDS'),
                  postcode = replace(postcode, achd_id == 4401, 2617),
                  suburb = replace(suburb, achd_id == 1614, 'INGLEBURN'),
                  suburb = replace(suburb, achd_id == 3531, 'HUNTERS HILL'),
                  suburb = replace(suburb, achd_id == 4350, 'QUEANBEYAN'),
                  valid_address = replace(valid_address, achd_id == 1194, 0),
                  valid_address = replace(valid_address, achd_id == 3288, 0),
                  suburb = replace(suburb, achd_id == 4638, 'REDFERN'),
                  suburb = replace(suburb, achd_id == 5100, 'GREENACRE'),
                  suburb = replace(suburb, achd_id == 1308, 'CURRANS HILL'),
                  postcode = replace(postcode, achd_id == 3827, 2008),
                  postcode = replace(postcode, achd_id == 4637, 2150),
                  suburb = replace(suburb, achd_id == 4502, 'WOOLLOOMOOLOO'),
                  postcode = replace(postcode, achd_id == 2850, 2360),
                  postcode = replace(postcode, achd_id == 2888, 2006),
                  postcode = replace(postcode, achd_id == 1643, 2129),
                  suburb = replace(suburb, achd_id == 3988, 'WARRIEWOOD'),
                  suburb = replace(suburb, achd_id == 4181, 'AMBARVALE'),
                  suburb = replace(suburb, achd_id == 1280, 'BONNELLS BAY'),
                  postcode = replace(postcode, achd_id == 1411, 2611),
                  postcode = replace(postcode, achd_id == 2680, 2130),
                  suburb = replace(suburb, achd_id == 3537, 'SOUTH GUNDAGAI'),
                  suburb = replace(suburb, achd_id == 2593, 'ARTARMON'),
                  postcode = replace(postcode, achd_id == 3503, 2750),
                  postcode = replace(postcode, achd_id == 2304, 2612),
                  suburb = replace(suburb, achd_id == 2304, 'CAMPBELL'),
                  postcode = replace(postcode, achd_id == 4266, 2006),
                  valid_address = replace(valid_address, achd_id == 3006, 0),
                  valid_address = replace(valid_address, achd_id == 48, 0),
                  postcode = replace(postcode, achd_id == 288, 2913),
                  suburb = replace(suburb, achd_id == 3961, 'GOULBURN'),
                  postcode = replace(postcode, achd_id == 2595, 2068),
                  postcode = replace(postcode, achd_id == 5253, 2095),
                  suburb = replace(suburb, achd_id == 800, 'WAVERLY'),
                  postcode = replace(postcode, achd_id == 1247, 2580),
                  suburb = replace(suburb, achd_id == 4580, 'AMBARVALE'),
                  postcode = replace(postcode, achd_id == 4416, 2077),
                  suburb = replace(suburb, achd_id == 3832, 'SILVERWATER'),
                  postcode = replace(postcode, achd_id == 1922, 2557),
                  postcode = replace(postcode, achd_id == 211, 2232),
                  postcode = replace(postcode, achd_id == 1771, 2006),
                  valid_address = replace(valid_address, achd_id == 4002, 0),
                  suburb = replace(suburb, achd_id == 1977, 'GREENACRE'),
                  suburb = replace(suburb, achd_id == 2330, 'ORANGE'),
                  postcode = replace(postcode, achd_id == 1311, 2151),
                  suburb = replace(suburb, achd_id == 4518, 'CHERRYBROOK'),
                  suburb = replace(suburb, achd_id == 4686, 'TINGIRA HEIGHTS'),
                  postcode = replace(postcode, achd_id == 2962, 2902),
                  postcode = replace(postcode, achd_id == 4161, 2516),
                  postcode = replace(postcode, achd_id == 4008, 2114),
                  suburb = replace(suburb, achd_id == 4349, 'TALBINGO'),
                  postcode = replace(postcode, achd_id == 2937, 2209),
                  valid_address = replace(valid_address, achd_id == 3096, 0),
                  postcode = replace(postcode, achd_id == 3802, 2560),
                  postcode = replace(postcode, achd_id == 5177, 2631),
                  suburb = replace(suburb, achd_id == 2645, 'WATSONS BAY'),
                  valid_address = replace(valid_address, achd_id == 2547, 0),
                  postcode = replace(postcode, achd_id == 3989, 2019),
                  postcode = replace(postcode, achd_id == 3599, 2207),
                  suburb = replace(suburb, achd_id == 4535, 'BLUE HAVEN'),
                  suburb = replace(suburb, achd_id == 4605, 'STRATHFIELD'),
                  suburb = replace(suburb, achd_id == 5037, 'SADLEIR'),
                  postcode = replace(postcode, achd_id == 4280, 2010),
                  suburb = replace(suburb, achd_id == 4544, 'WESTON'),
                  valid_address = replace(valid_address, achd_id == 4321, 0),
                  suburb = replace(suburb, achd_id == 4684, 'AMBARVALE'),
                  postcode = replace(postcode, achd_id == 1012, 2446),
                  postcode = replace(postcode, achd_id == 4656, 2196),
                  suburb = replace(suburb, achd_id == 4636, 'ERSKINEVILLE'),
                  suburb = replace(suburb, achd_id == 4575, 'LEICHHARDT'),
                  valid_address = replace(valid_address, achd_id == 4366, 0),
                  suburb = replace(suburb, achd_id == 4793, 'JINDABYNE'),
                  suburb = replace(suburb, achd_id == 4659, 'FAIRFIELD HEIGHTS'),
                  suburb = replace(suburb, achd_id == 4472, 'STRATHFIELD'),
                  postcode = replace(postcode, achd_id == 4452, 2067),
                  valid_address = replace(valid_address, achd_id == 4422, 0),
                  suburb = replace(suburb, achd_id == 4618, 'ANNANGROVE'),
                  suburb = replace(suburb, achd_id == 5024, 'BANKSTOWN'),
                  valid_address = replace(valid_address, achd_id == 4567, 0),
                  suburb = replace(suburb, achd_id == 5024, 'SYLVANIA'),
                  postcode = replace(postcode, achd_id == 4453, 2067),
                  postcode = replace(postcode, achd_id == 4327, 2060),
                  suburb = replace(suburb, achd_id == 4803, 'BIRCHGROVE'),
                  suburb = replace(suburb, achd_id == 5063, 'NICHOLLS'),
                  suburb = replace(suburb, achd_id == 4783, 'TARRAWANNA'),
                  suburb = replace(suburb, achd_id == 4924, 'FRENCHS FOREST'),
                  postcode = replace(postcode, achd_id == 4927, 2067),
                  suburb = replace(suburb, achd_id == 4955, 'FAIRLIGHT'),
                  suburb = replace(suburb, achd_id == 5273, 'ST HELENS PARK'),
                  valid_address = replace(valid_address, achd_id == 240, 0),
                  postcode = replace(postcode, achd_id == 1538, 2006),
                  postcode = replace(postcode, achd_id == 876, 2753),
                  suburb = replace(suburb, achd_id == 1189, 'LONDONDERRY'),
                  postcode = replace(postcode, achd_id == 1559, 2474),
                  postcode = replace(postcode, achd_id == 4179, 2043),
                  suburb = replace(suburb, achd_id == 1091, 'WOLLONGONG WEST'),
                  suburb = replace(suburb, achd_id == 3503, 'EMU PLAINS'),
                  postcode = replace(postcode, achd_id == 5024, 2224),
                  suburb = replace(suburb, achd_id == 4655, 'SYLVANIA'),
                )
```



_MAIN_
``` {r}
# Load the MAIN table from the ACHD database
MAIN <- read.csv(paste(ACHD_path, 'main_Table.csv', sep="")) %>%
#       ------------Clean up the data-------------------------  
        # selecting non-dx variables only
        select(person_id, SHC, DC, adm_RC, adm_Bethesda, adm_Bethesda_Code, FupStatus, Death.Cause, achd.1) %>%
         rename('achd_id' = person_id,
               'dx_comment' = achd.1,
               'RC' = adm_RC,
               'bethesda_name' = adm_Bethesda,
               'bethesda_code' = adm_Bethesda_Code,
               'death' = FupStatus)
```



``` {r}
# Load the MAIN table from the ACHD database
DX <- read.csv(paste(ACHD_path, 'main_Table.csv', sep=""))%>%
#       ------------Clean up the data-------------------------  
        # selecting dx variables only
        select(person_id, ABNAO, AbnAV, AbnMV, ABNPV, AbnRV, ABNTV, adm_AbnCA, adm_AbnVeins, adm_AbsentPA, adm_ALCAPA, 
               adm_Aneurism, adm_APwindow, adm_AVfistula, adm_CHB, adm_CorTriatriatum, adm_Gerbodde, 
               adm_Hemi_truncus,adm_LPA_sling, adm_PAV_Malform, adm_Shones, AoC, AoInterrupt, AR, AS, ASD, 
               AVSD, BAVD, ccTGA, DEXTROCARDIA, DILV, DIRV, DORV, Ebstein.s, EISENMENGER.S, FONTAN, HLHS, 
               LVNONCOMPACT, MAPCA.s, Other, PA, PAPVD, PAS, PDA, PFO, PS, SubAS, SubPS, SupraAS, SupraMS,
               SupraPS, TA, TAPVD, TGA, TOF, TRUNCART, UnkDx, VSD) %>%
        rename('achd_id' = person_id) %>%
        # PA and and adm_AbsentPA are the same thing (Pulmonary Atresia or Absent Pulmonary Artery)
        # combine these into the same variable
        mutate(combine_PA = ifelse( (adm_AbsentPA != 0 | PA != 0 ), 1, 0)) %>%
        select(-adm_AbsentPA, -PA)

# Import the lookup table for EPCC codes, remove the old PA diagnoses
ACHD_to_EPCC <- read.csv('./input/ACHD_EPCC_coding.csv') %>% 
                filter(!(Variable == 'adm_AbsentPA' | Variable == 'PA')) %>%
                mutate(Variable = if_else(Variable == "Ebsteins", "Ebstein.s", Variable),
                       Variable = if_else(Variable == "EISENMENGERS", "EISENMENGER.S", Variable),
                       Variable = if_else(Variable == "MAPCAs", "MAPCA.s", Variable))

# Change the DX column names to the correct ECC codes
names(DX) <- ACHD_to_EPCC$EPCC_Code[match(names(DX), ACHD_to_EPCC$Variable)]

# Remove the 'achd_id' row from the ACHD to EPCC lookup table
ACHD_to_EPCC <- filter(ACHD_to_EPCC, !(Variable == "achd_id"))
```

All of the diagnosis variables code 1 or -1 for true and 0 or NA for untrue. To make it a little easier to read, we will recode -1 to 1 and NA to 0.

The death variable (FupStatus) also codes alive as NA or 0, recode all NA's to 0

``` {r}
# all the diagnosis variables
EPCC_codes <- as.character(ACHD_to_EPCC[["EPCC_Code"]])
dx_names <- as.character(ACHD_to_EPCC[["Adjust_name"]])

# recode the diagnoses to binary
DX <- DX %>% mutate_at(EPCC_codes,
                            function(x) replace(x, is.na(x), 0) # recode all na values to 0
                           ) %>%
                 mutate_at(EPCC_codes,
                            function(x) replace(x, x == -1, 1) # recode all -1 values to 1
                           ) %>%
                 mutate(no_dx = rowSums(.[,EPCC_codes]))

# Recode the death variable to binary
MAIN <- MAIN %>% mutate_at('death', 
                           function(x) replace(x, is.na(x), 0)) # recode all na values to 0)
```

in the disease severity code 'bethesda_code'; 4 codes for 'Classification Unknown', there are a few other codes (0,9,NA) that can be cleaned to 4 also

``` {r}
# recode the classification to unknown
MAIN <- MAIN %>% mutate_at('bethesda_code',
                            function(x) replace(x, (is.na(x) | x == 0 | x == 9), 4) # recode all na values to 0
                           )
```

Adding label to the EPCC codes. The EPCC codes aren't very readable, so we will add some labels that outline the diangosis names more clearly.

``` {r}
# Create a named list to use for adding variable labels
EPCC_to_names <- setNames(dx_names, EPCC_codes)

# Apply label to the diagnosis variables
library(expss) #for apply_labels
DX <- do.call(apply_labels, c(list(data = DX), EPCC_to_names))
```


_DEATH_
``` {r}
# load the DEATH table from the ACHD database
DEATH <- read.csv(paste(ACHD_path, 'DEATH.csv', sep="")) %>%
#        ------------Clean up the data-------------------------  
         # remove unnecessary columns
         select(person_id, death_date) %>%                             
         # remove timestamp from death date
         mutate_at("death_date", str_remove, pattern = " 0:00:00") %>%        
         # convert death date to date type
         mutate_at("death_date", as.Date, format='%d/%m/%Y') %>%              
         # rename columns
         rename("achd_id" = person_id)                                        
```

_ASGS AREAS_
``` {r}
# Import ASGS Lookup table
asgs_lookup <- read.csv("./output/ASGS_lookup.csv") %>%
#             ------------Clean up the data-----------------  
              # select subrub, postcode and sa2 name columns
              select(LOCALITY_NAME,POSTCODE,SA2_NAME, SA3_NAME, SA4_NAME, GCC_NAME) %>% 
              # Convert postcode to character
              mutate_at("POSTCODE", as.character) %>%    
              # Convert suburb name to character
              mutate_at("LOCALITY_NAME", as.character)    
```

``` {r}
# Add the ASGS area names to the ADMIN dataset, based on the suburb and postcode name
ADMIN_sa2 <- left_join(ADMIN, asgs_lookup, by = c("suburb" = "LOCALITY_NAME", 
                                                 "postcode" = "POSTCODE"))
```

_FOLLOW UP STATUS_
``` {r}
# load Follow Up Status dataset
FUP <- read.csv(paste(ACHD_path, 'FOLLOW-UP STATUS.csv', sep=""))  %>%
#               ------------Clean up the data-------------------------
                # select columns for id and clinic dates
                select(person_id, fup_last_date, fup_comment, fup_Seen_by, fup_outreach, fup_new) %>%
                # rename columns
                rename( 'achd_id' = person_id,                        
                        'clinic_date' = fup_last_date, 
                        'clinic_comment' = fup_comment,
                        'seen_by' = fup_Seen_by,
                        'outreach' = fup_outreach,
                        'new' = fup_new) %>%
                # remove timestamp from clinic date
                mutate_at("clinic_date", str_remove, pattern = " 0:00:00") %>%       
                # Convert clinic dates to Date format
                mutate_at("clinic_date",as.Date, format='%d/%m/%Y') %>%  
                # Arrange by clinic date in ascending order
                arrange(clinic_date) %>%
                # Filter for records without a missing clinic date
                filter(!is.na(clinic_date)) %>%
                # Replace missing in "seen_by" with NA
                mutate(seen_by = replace(seen_by, seen_by == "", NA))
```


``` {r}
# Join the tables to create a person level dataset

                # Add data from MAIN Table
person.level <- left_join(ADMIN_sa2, MAIN, by = "achd_id") %>%
                rename("sa2" = SA2_NAME,
                       "sa3" = SA3_NAME,
                       "sa4" = SA4_NAME,
                       "gcc" = GCC_NAME) %>%
                # Add data from the DX Table
                left_join(DX, by = "achd_id") %>%
                # Add data from DEATH table
                left_join(DEATH, by = 'achd_id') %>%
                # Add 0 flag for people still alive in 'death' variable
                mutate(death = replace(death, is.na(death), 0)) %>%
                # add age at death   
                mutate(age_death = as.period(dob %--% death_date)) %>% 
                # add data from Follow Up Status table
                nest_join(FUP, by = 'achd_id') %>%
                rename( 'clinics' =  FUP)
```

The dataframes contained in `clinics` column looks like this (for one record):

``` {r}
# display one of the clinic_dates dataframes
person.level %>% filter(achd_id == 1101) %>% .$clinics
```


To get some more information about the clinic visits for each patient, we will pull some information out of the `clinics` column

```{r warning = FALSE}
# Add some more rows to the dataset
person.level <- person.level %>%
            mutate(
              # Number of clinics for each patient
              no_clinics = map_dbl(map(clinics, ~ .$clinic_date), length), 
              # First clinic date for each patient
              first_clinic = as_date(map_dbl(clinics, ~ min(.$clinic_date))),
              # Last clinic date for each patient
              last_clinic = as_date(map_dbl(clinics, ~ max(.$clinic_date))), 
              # Number of days between first and last clinic
              date_diff = as.duration(first_clinic %--% last_clinic) / dyears(1),
              # age at first clinic attendance
              age_first_clinic = as.period(dob %--% first_clinic),
              # age at last clinic attendance
              age_last_clinic = as.period(dob %--% last_clinic),
              # time since last clinic attendance
              gap = case_when(
                death == 0 ~ as.duration(as.Date('31/12/2020', format='%d/%m/%Y') - last_clinic),
                death == 1 ~ as.duration(death_date - last_clinic),
               ))
```

``` {r warning = FALSE}
person.level <- person.level %>%
            mutate(
              # Remove visits before 01/01/2000
              clinics_2000 = map(clinics, ~ filter(., clinic_date >= as.Date('01/01/2000', format='%d/%m/%Y'))),
              # number of clinics for each patient
              no_clinics_2000 = map_dbl(map(clinics_2000, ~ .$clinic_date), length),
              # Flag to for patient with no visit before 2000
              clinic_drop = ifelse(no_clinics_2000 == 0, 1, 0),
              # First clinic date for each patient
              first_clinic_2000 = as_date(map_dbl(clinics_2000, ~ min(.$clinic_date))), 
              # Last clinic date for each patient
              last_clinic_2000 = as_date(map_dbl(clinics_2000, ~ max(.$clinic_date))),  
              # Years between first and last clinic 
              date_diff_2000 = as.duration(first_clinic_2000%--%last_clinic_2000)/dyears(1),   
              # Recode infinity to NA
              date_diff_2000 = replace(date_diff_2000, date_diff_2000 == -Inf, NA),
              # age at first clinic attendance
              age_first_clinic_2000 = as.period(dob %--% first_clinic_2000),
              # age at last clinic attendance
              age_last_clinic_2000 = as.period(dob %--% last_clinic_2000),
              # time since last clinic attendance
              gap_2000 = case_when(
                death == 0 ~ as.duration(as.Date('01/07/2022', format='%d/%m/%Y') - last_clinic_2000),
                death == 1 ~ as.duration(death_date - last_clinic_2000),
               ),
              # not seen for more than 3 years
              ltf_3 = ifelse(as.numeric(gap_2000, 'years') > 3, 1, 0),
              # not seen for more than 4 years
              ltf_4 = ifelse(as.numeric(gap_2000, 'years') > 4, 1, 0),
              # not seen for more than 5 years
              ltf_5 = ifelse(as.numeric(gap_2000, 'years') > 5, 1, 0),
                   )
```

__Address Cleaning List__
Final version of this document should have nothing in it, address cleaning performed above was used with the cleaning list dated 08/10/2020
``` {r}
address_cleaning_list <- person.level %>% filter(valid_address == 1) %>%
                                           filter(is.na(sa2)) %>%
                                           select(achd_id,postcode, suburb, state,)

address_cleaning_list

#write.csv(address_cleaning_list, paste("Z:/CURRENT_STUDIES/2020_achd_map/DATA/", 'output/test_', Sys.Date(), '.csv', sep=""))
```

__Cleaning Data Types__
``` {r, echo = FALSE}
# Variables to convert to factor
names_factor <- c('achd_id' ,'sex', 'age_flag', 'RC', 'death', 'clinic_drop', 'valid_address', 'bethesda_name', 
                  'bethesda_code', EPCC_codes)

# Convert the above variables to factors
person.level[,names_factor] <- lapply(person.level[,names_factor] , factor)

# convert the diagnosis comment to a character vector
person.level$dx_comment <- as.character(person.level$dx_comment)

#re-add the diagnosis labels
person.level <- do.call(apply_labels, c(list(data = person.level), EPCC_to_names))
```


``` {r}
print('Variables and data types')
sapply(person.level, class)
```

__EDA__  
The data cleaning above has created three variables that we will use to filter out invalid records  
    - age_flag: flags people 18 and over (1) or below 18 (0)  
    - valid_address: flags people with valid addresses (1) or invalid (0)  
    - clinic_drop: flags records with no clinic visits after 01/01/2000 (1) or those with clinics after that date (0)  

The purpose of this EDA is to explore how filtering out records based on the criteria above affects important data

_Distribution of patient above and below 18 years old_
``` {r, echo = FALSE, warning = FALSE}
knitr::kable(person.level %>% count(age_flag)) 
```

_Distribution of patient with and without clinic dates after 01/01/2000_
``` {r, echo = FALSE, warning = FALSE}
knitr::kable(person.level %>% count(clinic_drop))
```

_Distribution of patients with and without valid addresses_
``` {r, echo = FALSE, warning = FALSE}
knitr::kable(person.level %>% count(valid_address))
```

Create a second dataset that filters out the above  
    - Patients 18 or over (`age_flag == 1`)  
    - Patients with a valid address (`valid_address == 1`)  
    - Patients with clinic visits after 01/01/2020 (`clinic_drop == 0`)
    
``` {r}
person.level.filtered <- person.level %>%
                            filter(age_flag == 1 &
                                   valid_address == 1 &
                                   clinic_drop == 0)

```


``` {r, echo = FALSE}
print("Dimensions of the original dataset")
paste('Number of variables:', dim(person.level)[2])
paste('Number of records:', dim(person.level)[1])

print("Dimensions of the filtered dataset")
paste('Number of variables:', dim(person.level.filtered)[2])
paste('Number of records:', dim(person.level.filtered)[1])
```


``` {r}
# check missingness in the original dataset
sapply(person.level, function(x) sum(is.na(x)))

```

``` {r}
# check missingness in the filtered dataset
sapply(person.level.filtered, function(x) sum(is.na(x)))

```

_Number of Clinics_
``` {r, echo = FALSE}
ggplot(person.level, aes(x=no_clinics)) +
      geom_bar(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.7), labels = percent) +
      scale_x_continuous(breaks = seq(0,20,1)) +
      labs(title = "Distribution of Number of Clinic Visits", 
           y = "Percent", 
           x = "Number of Visits")


ggplot(person.level.filtered, aes(x=no_clinics_2000)) +
      geom_bar(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.7), labels = percent) +
      scale_x_continuous(breaks = seq(0,20,1)) +
      labs(title = "Distribution of Number of Clinic Visits after 01/01/2000", 
           y = "Percent", 
           x = "Number of Visits")

```

``` {r, echo = FALSE}
person.level %>%
  filter(no_clinics > 1) %>%
  ggplot(aes(x=no_clinics)) +
      geom_bar(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.5), labels = percent) +
      scale_x_continuous(breaks = seq(0,20,1)) +
      labs(title = "Distribution of Number of Clinic Visits", 
           y = "Percent", 
           x = "Number of Visits")

person.level.filtered %>%
  filter(no_clinics_2000 > 1) %>%
  ggplot(aes(x=no_clinics_2000)) +
        geom_bar(aes( y = ..count../sum(..count..))) +
        scale_y_continuous(limits = c(0,0.5), labels = percent) +
        scale_x_continuous(breaks = seq(0,20,1)) +
        labs(title = "Distribution of Number of Clinic Visits after 01/01/2000", 
             y = "Percent", 
             x = "Number of Visits")

```

_First Clinic Date_
``` {r, echo = FALSE}
ggplot(person.level, aes(x=first_clinic)) +
      geom_histogram(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.25), labels = percent) +
      labs(title = "Distribution of First Clinic Visit", 
           y = "Percent", 
           x = "Year")

ggplot(person.level.filtered, aes(x=first_clinic_2000)) +
      geom_histogram(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.25), labels = percent) +
      labs(title = "Distribution of First Clinic Visit after 01/01/2000", 
           y = "Percent", 
           x = "Year")
```

_Last Clinic Date_
``` {r, echo = FALSE}
ggplot(person.level, aes(x=last_clinic)) +
      geom_histogram(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.25), labels = percent) +
      labs(title = "Distribution of Last Clinic Visit", 
           y = "Percent", 
           x = "Year")

ggplot(person.level.filtered, aes(x=last_clinic_2000)) +
      geom_histogram(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.25), labels = percent) +
      labs(title = "Distribution of Last Clinic Visit after 01/01/2000", 
           y = "Percent", 
           x = "Year")
```

_Time between first and last clinic_
``` {r, echo = FALSE}
ggplot(person.level, aes(x=date_diff)) +
      geom_histogram(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.7), labels = percent) +
      labs(title = "Distribution of Years between first and last visit", 
           y = "Percent", 
           x = "years")

ggplot(person.level.filtered, aes(x=date_diff_2000)) +
      geom_histogram(aes( y = ..count../sum(..count..))) +
      scale_y_continuous(limits = c(0,0.7), labels = percent) +
      labs(title = "Distribution of Years between first and last visit, after 01/01/2000", 
           y = "Percent", 
           x = "years")
```

``` {r, echo = FALSE}
person.level %>%
  filter(no_clinics > 1) %>%
  ggplot(aes(x=date_diff)) +
      geom_histogram(aes( y = ..count../sum(..count..)))+
      scale_y_continuous(limits = c(0,0.3), labels = percent) +
      labs(title = "Distribution of Years between first and last visit", 
           y = "Percent", 
           x = "years")

person.level.filtered %>%
  filter(no_clinics_2000 > 1) %>%
  ggplot(aes(x=date_diff_2000)) +
        geom_histogram(aes( y = ..count../sum(..count..)))+
        scale_y_continuous(limits = c(0,0.3), labels = percent) +
        labs(title = "Distribution of Years between first and last visit, after 01/01/2000", 
             y = "Percent", 
             x = "years")
```

_sex_
``` {r, echo = FALSE}
ggplot(person.level, aes(x=as.factor(sex))) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,0.6), breaks = seq(0,0.6,0.1), labels = percent) +
  labs(title = "Sex", 
       y = "Percent", 
       x = "Sex")


ggplot(person.level.filtered, aes(x=as.factor(sex))) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,0.6), breaks = seq(0,0.6,0.1), labels = percent) +
  labs(title = "Sex", 
       y = "Percent", 
       x = "Sex")
```

_age_
```{r, echo = FALSE}
person.level %>%
  filter(death == 0) %>%
  ggplot(aes(x=as.numeric(age, "years"))) +
    geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
    scale_y_continuous(limits = c(0,0.1), labels = percent) +
    scale_x_continuous(limits = c(0,120)) +
    labs(title = "Age", 
         y = "count", 
         x = "years")

person.level.filtered %>%
  filter(death == 0) %>%
  ggplot(aes(x=as.numeric(age, "years"))) +
    geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
    scale_y_continuous(limits = c(0,0.1), labels = percent) +
    scale_x_continuous(limits = c(0,120)) +
    labs(title = "Age", 
         y = "count", 
         x = "years")


```

_death_
```{r, echo = FALSE}
ggplot(person.level, aes(x=as.factor(death))) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,1), labels = percent) +
  labs(title = "death", 
       y = "Percent", 
       x = "death")

ggplot(person.level.filtered, aes(x=as.factor(death))) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,1), labels = percent) +
  labs(title = "death", 
       y = "Percent", 
       x = "death")
```

_age at death_
```{r, echo = FALSE}
ggplot(person.level, aes(x=as.numeric(age_death, "years"))) +
  geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
  scale_y_continuous(limits = c(0,0.1), labels = percent) +
  labs(title = "age at death", 
       y = "count", 
       x = "age at death")

ggplot(person.level.filtered, aes(x=as.numeric(age_death, "years"))) +
  geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
  scale_y_continuous(limits = c(0,0.1), labels = percent) +
  labs(title = "age at death", 
       y = "count", 
       x = "age at death")
```


_Diagnoses_
``` {r echo = FALSE}
ggplot(person.level, aes(x=no_dx)) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,1), labels = percent) +
  scale_x_continuous(breaks = seq(0,10,1)) +
  labs(title = "Number of Diangoses per person", 
       y = "count", 
       x = "Number of Diagnoses")

ggplot(person.level.filtered, aes(x=no_dx)) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,1), labels = percent) +
  scale_x_continuous(breaks = seq(0,10,1)) +
  labs(title = "Number of Diangoses per person", 
       y = "count", 
       x = "Number of Diagnoses")
```

``` {r echo = FALSE}
print('number of people with no diagnosis - original dataset:')
length(person.level$achd_id[person.level$no_dx == 0])

print('number of people with no diagnosis - filtered dataset:')
length(person.level.filtered$achd_id[person.level.filtered$no_dx == 0])
```

``` {r, echo = FALSE}
# created a dataset with only the disagnoses
dx.only <- person.level %>%
  select(all_of(EPCC_codes)) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

#re-add the diagnosis labels
dx.only <- do.call(apply_labels, c(list(data = dx.only), EPCC_to_names))

# created a dataset with only the disagnoses - from the filtered data
dx.only.filtered <- person.level.filtered %>%
  select(all_of(EPCC_codes)) %>%
  mutate_all(as.character) %>%
  mutate_all(as.numeric)

#re-add the diagnosis labels
dx.only.filtered <- do.call(apply_labels, c(list(data = dx.only.filtered), EPCC_to_names))
```

``` {r, echo = FALSE}
# The frequency each diagnosis appears in the dataset
dx.count <- dx.only %>%
            summarise_all(sum, na.rm = TRUE) %>%
            t() 

# Convert to data frame and clean
dx.count <- as.data.frame(dx.count)
dx.count$dx_name <- dx_names

# plot the frequency of each diagnosis
ggplot(dx.count, aes(x = dx_name, y=V1)) + 
geom_bar(stat="identity",) +
theme(axis.text.x = element_text(angle = 90)) +
labs(title = "Frequecy of each diagnosis - full dataset", 
   y = "count",
   x = "")
```


``` {r, echo = FALSE}
# The frequency each diagnosis appears in the dataset - from the filtered dataset
dx.count.filtered <- dx.only.filtered %>%
            summarise_all(sum, na.rm = TRUE) %>%
            t() 

# Convert to data frame and clean
dx.count.filtered <- as.data.frame(dx.count.filtered)
dx.count.filtered$dx_name <- dx_names

# plot the frequency of each diagnosis
ggplot(dx.count.filtered, aes(x = dx_name, y=V1)) + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Frequecy of each diagnosis - filtered dataset", 
     y = "count",
     x = "")
```

_Disease Severity_
``` {r}
ggplot(person.level, aes(x=bethesda_code)) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,1), labels = percent) +
  labs(title = "disease severity", 
       y = "Percent", 
       x = "death")

ggplot(person.level.filtered, aes(x=bethesda_code)) +
  geom_bar(aes( y = ..count../sum(..count..)))+
  scale_y_continuous(limits = c(0,1), labels = percent) +
  labs(title = "disease severity", 
       y = "Percent", 
       x = "death")
```



_Counting the ACHD population in each area_
``` {r}
# count the number achd patients in each SA2 area
sa2.count <- person.level.filtered %>% count(sa2) %>%
                        rename("ACHD_count" = n)

# count the number achd patients in each SA3 area
sa3.count <- person.level.filtered %>% count(sa3) %>%
                        rename("ACHD_count" = n)

# count the number achd patients in each SA4 area
sa4.count <- person.level.filtered %>% count(sa4) %>%
                        rename("ACHD_count" = n)
```


``` {r}
sa2.count %>% filter(!is.na(sa2)) %>%
ggplot(aes(x=ACHD_count)) +
  geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
  scale_y_continuous(limits = c(0,0.25), labels = percent) +
  scale_x_continuous(breaks = seq(0,60,5)) +
  labs(title = "Population per sa2 area", 
       y = "percent", 
       x = "Population per sa2 area")

sa3.count %>% filter(!is.na(sa3)) %>%
ggplot(aes(x=ACHD_count)) +
  geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
  scale_y_continuous(limits = c(0,0.25), labels = percent) +
  scale_x_continuous(breaks = seq(0,500,50)) +
  labs(title = "Population per sa3 area", 
       y = "percent", 
       x = "Population per sa3 area")

sa4.count %>% filter(!is.na(sa4)) %>%
ggplot(aes(x=ACHD_count)) +
  geom_histogram(aes( y = ..count../sum(..count..)), bins = 50)+
  scale_y_continuous(limits = c(0,0.25), labels = percent) +
  scale_x_continuous(breaks = seq(0,500,50)) +
  labs(title = "Population per sa4 area", 
       y = "percent", 
       x = "Population per sa4 area")

```

_Saving the data_
Will not run by default to not overwrite data accidentally, change eval = if you need to rewrite the patient data
``` {r, eval = TRUE}
#Save the Master Code Sheet
# request a password to protect the master code sheet
# mcs_pass <- readline(prompt="enter a password for the mastercode sheet: ")
# select the identifiers from the dataset
master_code_sheet <- person.level.filtered %>% select(achd_id, dob, sex, suburb, postcode, state)
# write the master code sheet to an excel spreadsheet
write.csv(x = master_code_sheet,
           file = paste(ACHD_path, "RPAH_master_code_sheet", Sys.Date(), ".csv", sep=""))

#Save the analysis data set
analysis_dataset <- person.level.filtered %>% select(-dob, -suburb, -postcode, -state,
                                                     -clinics, -no_clinics, -first_clinic, -last_clinic,
                                                     -date_diff, -age_first_clinic, -age_last_clinic, -gap)
#write the analysis dataset to a csv file and an R data file
saveRDS(analysis_dataset, file = paste(ACHD_path, "output/rpah_analysis_dataset.rds", sep=""))
#cannot save the nested 'clinics columns as csv
analysis_dataset %>% select(-clinics_2000) %>%
      write.csv(paste(ACHD_path, 'output/rpah_analysis_dataset.csv', sep=""))
```

``` {r}
# Some summary numbers for publication
paste("age mean: ", mean(as.numeric(analysis_dataset$age, 'years')))
paste("age sd: ", sd(as.numeric(analysis_dataset$age, 'years')))

paste("female: ", ( analysis_dataset %>% filter(sex == 2) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("neither m or f: ", ( analysis_dataset %>% filter(sex == 4) %>% nrow() ))

paste("Bicuspid: ", ( analysis_dataset %>% filter(`09.15.22` == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("VSD: ", ( analysis_dataset %>% filter(`07.10.00` == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("ASD: ", ( analysis_dataset %>% filter(`05.03.00` == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("TOF: ", ( analysis_dataset %>% filter(`01.01.01` == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("PFO: ", ( analysis_dataset %>% filter(`05.03.01` == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )

paste("simple: ", ( analysis_dataset %>% filter(bethesda_code == 1) %>% nrow() ))
paste("moderate: ", ( analysis_dataset %>% filter(bethesda_code == 2) %>% nrow() ))
paste("complex: ", ( analysis_dataset %>% filter(bethesda_code == 3) %>% nrow() ))
paste("unknown: ", ( analysis_dataset %>% filter(bethesda_code == 4) %>% nrow() ))

paste("no clinics mean: ", mean(analysis_dataset$no_clinics_2000))
paste("no clinincs sd: ", sd(analysis_dataset$no_clinics_2000))

paste("gap mean: ", mean(as.numeric(analysis_dataset$gap_2000, 'years'), na.rm = TRUE))
paste("gap sd: ", sd(as.numeric(analysis_dataset$gap_2000, 'years'), na.rm = TRUE))

paste("ltf all: ", ( analysis_dataset %>% filter(ltf_3 == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("ltf simple: ", ( analysis_dataset %>% filter(bethesda_code == 1) %>% filter(ltf_3 == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("ltf moderate: ", ( analysis_dataset %>% filter(bethesda_code == 2) %>% filter(ltf_3 == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
paste("ltf complex: ", ( analysis_dataset %>% filter(bethesda_code == 3) %>% filter(ltf_3 == 1) %>% nrow() ) / ( analysis_dataset %>% nrow() ) )
```

``` {r}
hist(as.numeric(analysis_dataset$gap_2000, 'years'))
```




