
``` {r, echo = FALSE}
selected_area <- drive.values$area_data %>% filter(SA2_5DIGIT == {{i}})
```

## Area Focus: `r selected_area$sa2_area` 

&nbsp; 

<div style= "float:right;
             position: relative;
             left: -220px;
             ">
``` {r, echo = FALSE}
leaflet() %>%
            addTiles() %>%
            # Add Polygons
            addPolygons(
                data = subset(drive.polys(), drive.polys()$NAME16 == selected_area$sa2_area),
                color = 'black',
                fill = 'white')
```
</div>


<div style= "line-height: 2.5;
             font-size: 15px;">
Number of ACHD Patients: `r selected_area$ACHD_count`  
Number with: Simple CHD: `r selected_area$beth_1` | Moderate CHD: `r selected_area$beth_2` | Complex CHD: `r selected_area$beth_3`  
Number of Patients lost to follow up: `r selected_area$ltf_3`  
Index of Relative Socio-economic Disadvantage (IRSD): `r selected_area$IRSD`  
Accessibility and Remoteness Index of Australia (percentage of population living in):  

* Major Cities: `r percent(selected_area$major_cities)`  
* Inner Regional: `r percent(selected_area$inner_regional)`  
* Outer Regional: `r percent(selected_area$outer_regional)`  
* Remote: `r percent(selected_area$remote)`  
* Very Remote: `r percent(selected_area$very_remote)`  

Driving time to nearest clinic:  

* Current Clinic Locations: `r round(as.numeric(area.achd$shortest_time[area.achd$sa2_area == selected_area$sa2_area], 'hours'), digits = 2)` hours
* With New Clinics Added: `r round(as.numeric(selected_area$shortest_time, 'hours'), digits = 2)` hours

</div>

